---
title: Visualizing ICC, Discriminability and Identification Rate of Functional Connectivity
author: "Virgilio Gonzenbach & Andrew Chen"
date: "3/22/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      cache.path = "cache/",
                      fig.path = 'figures/',
                      dpi = 300)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r load_data, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(circlize)
library(ggplot2)
library(ComplexHeatmap)

load.data = function(data.type){
  #' Load and prepares data from .csv files
  
  # Check arguments
  if (!(data.type %in% c("sparse", "dense"))) stop("Data type must be 'sparse' or 'dense'")
  
  df = read.csv(sprintf("data/mean_icc_%s.csv", data.type))
  colnames(df) = c("Distance", "ROI", "Session", "ICC")
  df$ROI = as.numeric(gsub("ROI", "", df$ROI))
  
  # Merge .csv files
  df = read.csv(sprintf("data/dis_idr_sub_%s.csv", data.type)) %>% merge(df, by = c("Distance", "ROI", "Session"), sort = FALSE)
  colnames(df) = gsub("Sub_", "", colnames(df))
  colnames(df)[5] = "Identification Rate"
  
  # Order factors
  df$Distance = recode(df$Distance, `sqeuclidean` = "Squared Euclidean", `partial_correlation` = "Partial Correlation")
  df$Distance = tools::toTitleCase(df$Distance)
  df$Distance = factor(df$Distance, levels = rev(c("Correlation", "Cosine", "Covariance", "Euclidean", "Squared Euclidean", "Tangent", "Partial Correlation")))
  df$ROI = factor(df$ROI, levels = sort(unique(df$ROI)))
  
  return(df)
}

dense_df = load.data("sparse")
sparse_df = load.data("dense")
```

```{r heatmap_funcs, warning=FALSE, message=FALSE}
make_heatmap = function(df, metric){
    #' Returns heatmap
    heat <- NULL 
    
    for (dist_df in split(df, df$Distance)){
      heat_sub = dist_df %>% 
        select(all_of(metric), "ROI", "Session") %>% 
        pivot_wider(names_from = Session, values_from = metric)
      heat = rbind(heat, as.matrix(heat_sub[,-1]))
    }
    rownames(heat) = rep(unique(df$ROI), times = length(levels(df$Distance)))
    return(heat)
}

plot_heatmap = function(df, metric, f, col.range = "fitted", n.col = 11, ends = NULL){
  
  heat = make_heatmap(df, metric)
  
  n.sectors = length(unique(f))
  circos.clear() # Try w/o this line
  circos.par(gap.after = c(rep(2, n.sectors - 1), 20))
  
  if(is.null(ends)){
    # ends = round(range(df[,metric]),1) # Try floor and ceiling
    ends = range(df[,metric])
    ends = c(floor(ends[1]*10)*.1, 
             ceiling(ends[2]*10)*.1)
  }
  
  base_colors = RColorBrewer::brewer.pal(9, 'Spectral')
  names(base_colors) = seq(.2, 1, length.out=9)
  breaks = seq(ends[1], ends[2], .1)
  colors = base_colors[sapply(breaks, as.character)]
  col_fun = colorRamp2(breaks, colors, space="RGB")
  circos.heatmap(heat[,10:1], split = rep(levels(f), each = 10), col = col_fun,
                 track.height = 0.4,
                 cluster = FALSE,
                 rownames.side = "inside",
                 show.sector.labels = TRUE)
  
  circos.track(track.index = get.current.track.index(), panel.fun = function(x, y) {
    if(CELL_META$sector.numeric.index == n.sectors) { # the last sector
      cn = colnames(heat)
      n = length(cn)
      circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
                  1:n - 0.5, cn, 
                  cex = 0.5, adj = c(0, 0.5), facing = "inside")
      }
    }, bg.border = NA)
  
  lgd = Legend(title = metric, title_position = 'topcenter', col_fun = col_fun)
  grid.draw(lgd)
}
```

## Previsualization

### Ranges

```{r, results='asis'}
prt.rng = function(df){
  ranges = c("ICC", "Discriminability", "Identification Rate") %>% lapply(function(x){
      round(range(df[,x]), 2)
  })

  names(ranges) = c("ICC", "Discriminability", "Identification Rate")
  
  for (x in names(ranges)){
    cat(sprintf("- %s: %s - %s", x, ranges[[x]][1], ranges[[x]][2]), sep = "\n", fill=TRUE)
  }
}

cat("Dense: \n", fill=TRUE)
prt.rng(dense_df)
cat("\n Sparse: \n", fill=TRUE)
prt.rng(sparse_df)
```

### Histograms

Discriminability and Identification Rate values of 1 dominate the heatmaps.

```{r hist_dense}
dense_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identification Rate"), names_to="Metric", values_to = "Value") %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Dense FC Data")
```

```{r hist_dense_no1}
dense_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identification Rate"), names_to="Metric", values_to = "Value") %>% filter(Value < 1) %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Dense FC Data < 1")
```

```{r hist_sparse}
sparse_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identification Rate"), names_to="Metric", values_to = "Value") %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Sparse FC Data")
```

```{r hist_sparse_no1}
sparse_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identification Rate"), names_to="Metric", values_to = "Value") %>% filter(Value < 1) %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Sparse FC Data < 1")
```

## Circle Plots (Fitted color range)

Color endpoints of each circle plot were fixed to the minimum and maximum value (floored and ceilinged to the nearest tenth respectively) to represent encompass all of the data for a given heatmap.

### Dense

```{r dense_ICC_circle_fit, warning=FALSE, message=FALSE}
plot_heatmap(df = dense_df, metric = "ICC", f = dense_df$Distance)
```

```{r dense_discrim_circle_fit}
plot_heatmap(df = dense_df, metric = "Discriminability", f = dense_df$Distance)
```

```{r dense_identify_circle_fit}
plot_heatmap(df = dense_df, metric = "Identification Rate", f = dense_df$Distance)
```

### Sparse

```{r sparse_ICC_circle_fit, warning=FALSE, message=FALSE}
plot_heatmap(df = sparse_df, metric = "ICC", f = sparse_df$Distance)
```

```{r sparse_discrim_circle_fit}
plot_heatmap(df = sparse_df, metric = "Discriminability", f = sparse_df$Distance)
```

```{r sparse_identify_circe_fit}
plot_heatmap(df = sparse_df, metric = "Identification Rate", f = sparse_df$Distance)
```

## Circle Plots (Manual color range)

To make colors comparable between dense and sparse data, I fixed the color endpoints for each of the metrics:

-   ICC: 0.2 - 1
-   Discriminability: 0.8 - 1
-   Identification Rate: 0.6 - 1

### Dense

```{r dense_ICC_circle_manual}
plot_heatmap(df = dense_df, metric = "ICC", f = dense_df$Distance, ends=c(0.2, 0.8))
```

```{r dense_discrim_circle_manual}
plot_heatmap(df = dense_df, metric = "Discriminability", f = dense_df$Distance, ends=c(0.8, 1.0))
```

```{r dense_identify_circle_manual}
plot_heatmap(df = dense_df, metric = "Identification Rate", f = dense_df$Distance, ends=c(0.6, 1.0))
```

### Sparse

```{r sparse_ICC_circle_manual}
plot_heatmap(df = sparse_df, metric = "ICC", f = sparse_df$Distance, ends=c(0.2, 0.8))
```

```{r sparse_discrim_circle_manual}
plot_heatmap(df = sparse_df, metric = "Discriminability", f = sparse_df$Distance, ends=c(0.8, 1.0))
```

```{r sparse_identify_circle_manual}
plot_heatmap(df = sparse_df, metric = "Identification Rate", f = sparse_df$Distance, ends=c(0.6, 1.0))
```
