---
title: Visualizing ICC, Discriminability and Identifiability of Functional Connectivity

author: "Virgilio Gonzenbach & Andrew Chen"
date: "3/22/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = TRUE,
                      cache.path = "cache/",
                      fig.path = 'figures/',
                      dpi = 300)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r load_data, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(circlize)
library(ggplot2)
library(ComplexHeatmap)

load.data = function(data.type){
  #' Load and prepares data from .csv files
  
  # Check arguments
  if (!(data.type %in% c("sparse", "dense"))) stop("Data type must be 'sparse' or 'dense'")
  
  df = read.csv(sprintf("results/mean_icc_%s.csv", data.type))
  colnames(df) = c("Distance", "ROI", "Session", "ICC")
  df$ROI = as.numeric(gsub("ROI", "", df$ROI))
  
  # Merge .csv files
  df = read.csv(sprintf("results/dis_idr_sub_%s.csv", data.type)) %>% merge(df, by = c("Distance", "ROI", "Session"), sort = FALSE)
  colnames(df) = gsub("Sub_", "", colnames(df))
  colnames(df)[5] = "Identifiability"
  
  # Order factors
  df$Distance = factor(df$Distance, levels = rev(c("correlation", "cosine", "covariance", "euclidean", "sqeuclidean", "tangent", "partial_correlation")))
  df$ROI = factor(df$ROI, levels = sort(unique(df$ROI)))
  
  return(df)
}

dense_df = load.data("sparse")
sparse_df = load.data("dense")
```

```{r heatmap_funcs, warning=FALSE, message=FALSE}
make_heatmap = function(df, metric){
    #' Returns heatmap
    heat <- NULL 
    
    for (dist_df in split(df, df$Distance)){
      heat_sub = dist_df %>% 
        select(all_of(metric), "ROI", "Session") %>% 
        pivot_wider(names_from = Session, values_from = metric)
      heat = rbind(heat, as.matrix(heat_sub[,-1]))
    }
    rownames(heat) = rep(unique(df$ROI), times = length(levels(df$Distance)))
    return(heat)
}

plot_heatmap = function(df, metric, f, col.range = "fitted", n.col = 11, ends = NULL){
  
  heat = make_heatmap(df, metric)
  
  n.sectors = length(unique(f))
  circos.clear() # Try w/o this line
  circos.par(gap.after = c(rep(2, n.sectors - 1), 20))
  
  if(is.null(ends)){
    # ends = round(range(df[,metric]),1) # Try floor and ceiling
    ends = range(df[,metric])
    ends = c(floor(ends[1]*10)*.1, 
             ceiling(ends[2]*10)*.1)
  }
  
  col_fun =  colorRamp2(seq(ends[1], ends[2], length.out = n.col), RColorBrewer::brewer.pal(n.col, 'Spectral'), space="RGB")
  circos.heatmap(heat[,10:1], split = rep(unique(f), each = 10), col = col_fun,
                 track.height = 0.4,
                 cluster = FALSE,
                 rownames.side = "inside",
                 show.sector.labels = TRUE)
  
  circos.track(track.index = get.current.track.index(), panel.fun = function(x, y) {
    if(CELL_META$sector.numeric.index == n.sectors) { # the last sector
      cn = colnames(heat)
      n = length(cn)
      circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
                  1:n - 0.5, cn, 
                  cex = 0.5, adj = c(0, 0.5), facing = "inside")
      }
    }, bg.border = NA)
  
  lgd = Legend(title = metric, title_position = 'topcenter', col_fun = col_fun)
  grid.draw(lgd)
}
```

## Previsualization 

### Ranges

```{r, results='asis'}
prt.rng = function(df){
  ranges = c("ICC", "Discriminability", "Identifiability") %>% lapply(function(x){
      round(range(df[,x]), 2)
  })

  names(ranges) = c("ICC", "Discriminability", "Identifiability")
  
  for (x in names(ranges)){
    cat(sprintf("- %s: %s - %s", x, ranges[[x]][1], ranges[[x]][2]), sep = "\n", fill=TRUE)
  }
}

cat("Dense: \n", fill=TRUE)
prt.rng(dense_df)
cat("\n Sparse: \n", fill=TRUE)
prt.rng(sparse_df)
```

### Histograms
 
Discriminability and Identifiability values of 1 dominate the heatmaps.

```{r}
dense_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identifiability"), names_to="Metric", values_to = "Value") %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Dense FC Data")
```


```{r}
dense_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identifiability"), names_to="Metric", values_to = "Value") %>% filter(Value < 1) %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Dense FC Data < 1")
```

```{r}
sparse_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identifiability"), names_to="Metric", values_to = "Value") %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Sparse FC Data")
```

```{r}
sparse_df %>% pivot_longer(cols=c("ICC", "Discriminability", "Identifiability"), names_to="Metric", values_to = "Value") %>% filter(Value < 1) %>% ggplot(aes(x=Value)) + geom_histogram() + facet_wrap(~Metric) + ggtitle("Freq of values in Sparse FC Data < 1")
```

## Circle Plots (Fitted color range)

Color endpoints of each circle plot were fixed to the minimum and maximum value (floored and ceilinged to the nearest tenth respectively) to represent encompass all of the data for a given heatmap.

### Dense 

```{r dense_ICC_circle, warning=FALSE, message=FALSE}
plot_heatmap(df = dense_df, metric = "ICC", f = dense_df$Distance)
```

```{r dense_discrim_circle}
plot_heatmap(df = dense_df, metric = "Discriminability", f = dense_df$Distance)
```

```{r dense_identify_circle}
plot_heatmap(df = dense_df, metric = "Identifiability", f = dense_df$Distance)
```

### Sparse

```{r sparse_ICC_circle, warning=FALSE, message=FALSE}
plot_heatmap(df = sparse_df, metric = "ICC", f = sparse_df$Distance)
```

```{r sparse_discrim_circle}
plot_heatmap(df = sparse_df, metric = "Discriminability", f = sparse_df$Distance)
```

```{r sparse discrim}
plot_heatmap(df = sparse_df, metric = "Identifiability", f = sparse_df$Distance)
```

## Circle Plots (Manual color range)

To make colors comparable between dense and sparse data, I fixed the color endpoints for each of the metrics:

- ICC: 0.2 - 0.8
- Discriminability: 0.8 - 1
- Identifiability: 0.6 - 1


### Dense 

```{r dense_ICC_circle_0.2-0.8}
plot_heatmap(df = dense_df, metric = "ICC", f = dense_df$Distance, ends=c(0.2, 0.8))
```

```{r dense_discrim_circle_0.8-1}
plot_heatmap(df = dense_df, metric = "Discriminability", f = dense_df$Distance, ends=c(0.8, 1.0))
```

```{r dense_identify_circle_0.6-1}
plot_heatmap(df = dense_df, metric = "Identifiability", f = dense_df$Distance, ends=c(0.6, 1.0))
```

### Sparse

```{r sparse_ICC_circle_0.2-0.8}
plot_heatmap(df = sparse_df, metric = "ICC", f = sparse_df$Distance, ends=c(0.2, 0.8))
```

```{r sparse_discrim_circle_0.8-1}
plot_heatmap(df = sparse_df, metric = "Discriminability", f = sparse_df$Distance, ends=c(0.8, 1.0))
```

```{r sparse_identify_circle_0.6-1}
plot_heatmap(df = sparse_df, metric = "Identifiability", f = sparse_df$Distance, ends=c(0.6, 1.0))
```
